\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

This research was carried out at the Jet Propulsion Laboratory, California Institute of Technology, and was sponsored by JPL Visiting Student Research Program (JVSRP) and the National Aeronautics and Space Administration (80NM0018D0004).

I would like to extend my gratitude to the Jet Propulsion Laboratory, the Robotic Systems Lab, ETH Zurich, and the California Institute of Technology for making this collaborative work possible. I would like to thank my supervisors prof. Roland Brockers, Robert Hewitt PhD, and prof. Marco Hutter for their guidance and insights throughout these months.

It is with a feeling of overwhelming gratitude that I conclude this chapter and, with it, this thesis. The months involved with this work have been amongst the most educational of my whole life. I gathered tremendous amounts of knowledge and expertise from amazingly talented people and had uncountable memories that I will treasure forever. 

A few directed words of gratitude.

To Roland

For providing me with the opportunity to contribute my small part to NASA's inspiring journey. For all the guidance, insights and discussions which helped me push my boundaries and advance farther in this work. For a motivating nature to perform the best work possible and an honest and friendly spirit which made me feel at home.

To Luca

For being an inspiring bench neighbor and making me feel welcome from the very beginning.

To Dario 

For being an upbeat and honest nature and never failing to lighten a day.

To Simone

For a captivating honesty, and for showing me what is possible. 

To Rob and Pedro

For their expertise and their kind, helping hand.

To Adrian and Marco

For being a home - so far away from home and for never failing to inspire me to think outside the box.

To Micha, Alex, and Leona

For fun evenings and a refreshing presence.

To Mor and Far 

For giving me everything and expecting nothing in return.

To Ronja and Tove

For always listening and being there when needed.

To Jana

For never giving up and standing by my side.